---
title: "NYCFlights: Arrival Delay Logictic Model"
author: "Kevin Jue"
date: "5/2/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


## Logistic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }
## create the logistic function as defined in reading

## NOTE: created 2 functions, one which returns a single value of y between 0 and 1, and anothe that returns a vector of values for each vector of x, where each value is between 0 and 1

#function 1

logistic_func <- function(x) {
  y <-x
y <- 1/(1+prod(exp(-x)))
  
  return(y)
  }

## function 2

logistic_func2 <- function(x) {
  y <-x
  for(i in seq_along(x)){
y[i] <- 1/(1+prod(exp(-x[i])))
  }
  return(y)
  }


## test logistic function with some random numeric vector
x <- c(1,-2,3,4,-5,6,7,8,-9,-10)

logistic_func(x)

if(logistic_func(x) >= 0 & logistic_func(x) <= 1) {
  "okay"
} else {"bad"}


## plot the logistic function from [-10,10], ggplot2 was installed above

## note that we use the 2nd function as this returns a vector of elements all between 0 and 1

f <- ggplot(data.frame(x= c(-10,10)), aes(x))
f + stat_function(fun = logistic_func2, colour = "blue")

## inverse logistic function, solve arithmetically to find the inverse function of the logistic function of x
inverse_log <- function(x) {
  y <- x
  y <- log(x) - log(1-x)
  return (y)
}


## test inverse log function with some random numeric vector
x <- c(sample(0:100,10, replace = TRUE)/100)

inverse_log(x)

## we show that the range is -infinity to positive infinity by plotting it below
## ggplot2 was installed above

f <- ggplot(data.frame(x= c(0,1)), aes(x))
f + stat_function(fun = inverse_log, colour = "blue") + ylim(-50,50)



```


# NYCFlights Model

Using the rectangular data that you created from the earlier assignment and following theexample from the text and class, create a model for arr_delay >= 22 minutes. Describe/Explain each of the steps and show all work.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file to your repository.   

## create a logistic regression model that calculates the arr_delay >= 22 using a binomial/logistic classification model

## Step 1: Load Data
```{r "NYCFlights Logistic Regression Model" }

##reload the NYCFlights data from problem set 2, see below.  The final data set is called flights_join

#install the packages in order to complete the operation
library(readr)
library(magrittr)
library(dplyr)
library(data.table)
library(ggplot2)

#read the data into R
setwd('~/csx460/02-building-blocks/02-exercise-nycflights/data/')
flights <- setDT(read.csv('flights.csv'))
planes <- setDT(read.csv('planes.csv'))
airports <- setDT(read.csv('airports.csv'))
weather <- setDT(read.csv('weather.csv'))

#perform the necessary joins in order to make a rectangular data table

flights_join <- left_join(flights, planes, by = 'tailnum', suffix = c(".of_flights",".of_planes")) %>%
                left_join(weather, by = c('origin', 'time_hour'), suffix = c(".of_flights",".of_weather")) %>%
                left_join(airports, by = c('origin' = 'faa'), suffix = c(".of_flights",".of_origin")) %>%
                left_join(airports, by = c('dest' = 'faa'), suffix = c(".of_origin",".of_dest"))
```


## Step 2: Create Naive Model
```{r "Step 2: Set a baseline of model performance by calculating the the most commonly occuring value"}

## see exactly how many flights actually are delayed by 22 or more minutes
arr_delay_true <- sum(flights_join$arr_delay >= 22, na.rm = TRUE)

## see exactly how many flights are delayed by less than 22 minutes
arr_delay_false <- sum(flights_join$arr_delay < 22, na.rm = TRUE)

#odds are calculated p/q or p/(1-p)
odds_delay <- arr_delay_true/arr_delay_false

```



## Step 3: Calculate a baseline logistic regression
```{r "Step 3: As a first step run the simplest logistic regression with all variables as predictors"}

## change the response variable arr_delay to a factor to allow the logistic regression to be run


flights_join1 <- flights_join
flights_join1["arr_delay_22"] <- NA
for(i in seq(nrow(flights_join1))){
if(flights_join1$arr_delay[i] >= 22) {
  flights_join1$arr_delay_22[i] <- 1
} else {
    flights_join1$arr_delay_22[i] <- 0
  }
}



##convert to factor

flights_join1$arr_delay_22 <-  as.factor(flights_join1$arr_delay_22)

base_glm <- glm(formula = arr_delay_22 ~ year.of_flights + month.of_flights + day.of_flights + dep_time + sched_dep_time + dep_delay + arr_time + sched_arr_time + flight + air_time + distance + hour.of_flights + minute + year.of_planes + engines + seats + speed + year + month.of_weather + day.of_weather + hour.of_weather + temp + dewp + humid + wind_dir + wind_speed + wind_gust + precip + pressure + visib + lat.of_origin + lon.of_origin + alt.of_origin + tz.of_origin + lat.of_dest + lon.of_dest + alt.of_dest + tz.of_dest, family = binomial(link="logit"), data = flights_join1)

summary(base_glm)



```

# Question:

Is this a good model?  (Write your answer here.)


# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

