---
title: "Assignment NYC Flight Data"
author: "Kevin Jue"
date: "4/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. 
In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, 
`airports.csv`, `planes.csv` and `weather.csv`.  Join/merge these tables such 
that there is one rectangular table with one row for each flight. 


Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}
#install the packages in order to complete the operation
library(readr)
library(magrittr)
library(dplyr)
library(data.table)
library(ggplot2)


#read the csv file into r as a data.table
#below assumes that the csx460 folder and it's related subfolders are in the default working directory of your rstudio,
#if not then you'll need to change the WD (working directory)
setwd('~/csx460/02-building-blocks/02-exercise-nycflights/data/')
flights <- setDT(read.csv('flights.csv'))

#check that this is a data.table and a data.frame as well as view/analyze it's dimensions
str(flights)
head(flights)
```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}
#if you haven't loaded the packages in question 1, go and do so now
#create data.table with relevant data by selecting the columns we want, and then using melt to get them in long form
hist_da <- melt(setDT(data.frame(flights[,c(6,9)])))

#check that the data at least appears to have been translated okay!
head(flights$arr_delay) == head(hist_da[variable == 'arr_delay', value])
head(flights$dep_delay) == head(hist_da[variable == 'dep_delay', value])

#run the histogram using ggplot2
qplot(value, data=hist_da, fill = variable, geom = "histogram") + stat_bin(bins=40, binwidth = 50) +scale_x_continuous(breaks=seq(0,400,100))

```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}
#if you haven't loaded the packages from question 1, go and do so now
#not much to say, run a scatterplot in ggplot2, that's it
ggplot(flights[,c('day','arr_delay')], aes(x= day, y= arr_delay)) + geom_point(shape=1)
```



## Create tables for Categorical Variables

Create a `table`, counting the flights by origin airports and carrier.

```{r}
#if you haven't loaded the packages from question 1, go and do so now
#use dplyr:  group_by on origin and carrier (didn't see an airport so I assume that you were talking about the origin)
#this will change the two columns to factors, causing your summarise to sum along them.  use n() to count the rows.
flights_count <- flights %>% group_by(origin, carrier) %>% summarise(num_flights = n())

```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}
#if you haven't loaded the packages from question 1, go and do so now
#use dplyr and readr
setwd('~/csx460/02-building-blocks/02-exercise-nycflights/data/')
flights <- setDT(read.csv('flights.csv'))
planes <- setDT(read.csv('planes.csv'))
airports <- setDT(read.csv('airports.csv'))
weather <- setDT(read.csv('weather.csv'))

#Perform joins as appropriate, note that this will create alot of duplicate columns that are superfluous i.e.
#multiple day time and hour fields, so if i were to work with the data I would use a select() to get rid of them.
#However the assignment was to create a data table containing all columns from all tables with one record per row.
#so, that's what i did

flights_join <- left_join(flights, planes, by = 'tailnum', suffix = c(".of_flights",".of_planes")) %>%
                left_join(weather, by = c('origin', 'time_hour'), suffix = c(".of_flights",".of_weather")) %>%
                left_join(airports, by = c('origin' = 'faa'), suffix = c(".of_flights",".of_origin")) %>%
                left_join(airports, by = c('dest' = 'faa'), suffix = c(".of_origin",".of_dest"))
```

