---
title: "Caret / Recursive Partitioning"
author: "Kevin Jue"
date: "5/15/2017"
output: html_document
---

```{r init, warning=FALSE, echo=FALSE, message=FALSE}
library(rpart)
library(caret)
library(dplyr)
library(ggplot2)
library(data.table)
library(plotROC)
# .. Additional libraries
```


## Exercise 1: caret/logistic regression (5 points)

Rebuild your logistic regression model from the previous week, this time using the `caret` package. 

- Calculate the training or apparent performance of the model. 
- Calculate an unbiased measure of performance 
- Create a ROC Curve for your model

Show all work.

```{r logistic model}
#import flights data
setwd('~/csx460/02-building-blocks/02-exercise-nycflights/data/')
flights <- setDT(read.csv('flights.csv'))
planes <- setDT(read.csv('planes.csv'))
airports <- setDT(read.csv('airports.csv'))
weather <- setDT(read.csv('weather.csv'))

#perform the necessary joins in order to make a rectangular data table

flights_join <- left_join(flights, planes, by = 'tailnum', suffix = c(".of_flights",".of_planes")) %>%
                left_join(weather, by = c('origin', 'time_hour'), suffix = c(".of_flights",".of_weather")) %>%
                left_join(airports, by = c('origin' = 'faa'), suffix = c(".of_flights",".of_origin")) %>%
                left_join(airports, by = c('dest' = 'faa'), suffix = c(".of_origin",".of_dest"))

#use caret to run the same logistic regression that was run last week

#create the factor to run the logistic regression

flights_join1 <- flights_join
setDT(flights_join1)
flights_join1[, late := factor(arr_delay >= 22)]  # this is using the data.table package to create the factor


ctrl <- trainControl(method="boot",
                     #summaryFunction=twoClassSummary, 
                     #classProbs=TRUE,
                     savePredictions = TRUE)

 log_model <- train(late ~ dep_delay + arr_time + distance + month.of_weather + lat.of_dest + lon.of_dest, data = flights_join1, method = "glm", family = "binomial", trControl = ctrl, na.action = na.pass)

final_model <-log_model$finalModel
summary(final_model)

test <- log_model$pred

#calculate predictors

fit <- predict(final_model, flights_join1, type = 'response')
qplot(fit)

##calculate a confusionmatrix

logit.pred <- factor((fit) > 0.5, levels=c(TRUE, FALSE))
confusionMatrix(logit.pred, flights_join1$late)

# calculate an ROC curve

#couldn't get the ROC curve to work, it should look like this
#ggplot(log_model$pred$obs, 
#      aes(m = M, d = factor(obs, levels = c("R", "M")))) + 
#    geom_roc(hjust = -0.4, vjust = 1.5) + coord_equal()

```


## Exercise 2: caret/rpart (5 points)

Using the `caret` and `rpart` packages, create a **classification** model for flight delays using your NYC FLight data. Your solution should include:

- The use of `caret` and `rpart` to train a model.
- An articulation of the the problem your are 
- An naive model
- An unbiased calculation of the performance metric
- A plot of your model -- (the actual tree; there are several ways to do this)
- A discussion of your model 



Show and describe all work

```{r rpart model}
# Your Work Here

```


### Questions:

- Discuss the difference between the models and why you would use one model over the other?
- How might you produce an ROC type curve for the *rpart* model? 
