---
title: "Resampling"
author: "Kevin Jue"
date: "May 8, 2017"
output: html_document
---

**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file**

## Resampling

`x` is a random variable. We want to not only know what the `mean(x)` is but want to calculate the uncertainty of `mean(x)`.  Measuring the uncertainty requires repeated measurements of `mean(x)`.

- Calculate the mean of `x`.
- Calculte the `sd( mean(x) )` using the **using 10-fold resampling**.  Create your own folds, show your work.
- Calculte the `sd( mean(x) )` using the **using 10 bootstrap sample**.  Create your own folds, show your work. 

#create Functions
```{r}
## set seed and create a dataframe
set.seed(1984) #can be any number
k <- 10  # how many folds we want to create
n <- 40  # how large a sample size we want.  note that n/k should be an positive integer for this function to work
p <- 1   # lower bound
q <- 100 # upper bound

#call packages
library(magrittr)

x <- runif(n,p,q) %>% as.data.frame  #first variable is the num of obs, the other two give us the range to create

```

#calculate the mean of x
```{r}
#calculate the mean of x
x_mean <- mean(x[,])

```

#create the sd_cv function for n-fold resampling (in this case n=10)
```{r}
#create function that does n-fold resampling
sd_cv <- function(x) {
  y <- x
  #create appropriate folds in x
  folds <- as.data.frame(1:k)
  colnames(folds) <- c("folds")
  for(i in seq(1:k)){
    folds$folds[i] <- list(x[-((1+((nrow(x)/k)*(i-1))):((nrow(x)/k)*i)),])
  }
  # the above function maps each sampling set into each row of the data frame as an n-tupled list
  #now calculate the mean on each element of the dataframe (so the mean of the whole dataset) using lapply function
  #then calculate the SD
  folds1 <- as.data.frame(1:k)
  colnames(folds1) <- c("mean")
  for(j in seq(1:k)){
    folds1$mean[j] <- lapply(folds$folds[j], mean)
  }
  #convert/reduce from list to vector
  folds_sd <- as.numeric(unlist(folds1))
  SD_folds <- sd(folds_sd)
  return(SD_folds)
}
  
#test the function to see if it works

test_cv <- sd_cv(x)

```

#create sd_boot function for bootstrap resampling
```{r}
#create function that does bootstrap resampling
sd_boot <- function(x) {
  y <- x
  #create appropriate folds in x
  folds <- as.data.frame(1:k)
  colnames(folds) <- c("folds")
  for(i in seq(1:k)){
    folds$folds[i] <- list(x[sample(nrow(x),nrow(x), replace = TRUE),])
  }
  # the above function maps each sampling set into each row of the data frame as an n-tupled list
  #now calculate the mean on each element of the dataframe (so the mean of the whole dataset) using lapply function
  #then calculate the SD
  folds1 <- as.data.frame(1:k)
  colnames(folds1) <- c("mean")
  for(j in seq(1:k)){
    folds1$mean[j] <- lapply(folds$folds[j], mean)
  }
  #convert/reduce from list to vector
  folds_sd <- as.numeric(unlist(folds1))
  SD_folds <- sd(folds_sd)
  return(SD_folds)
}
#test the function to see if it works

test_boot <- sd_boot(x)

```


### Your answers here: 

### - sd_cv   is: 
```{r}
sd_cv
```
### - sd_boot is: 
```{r}
sd_boot
```