---
title: "NYCFlights: Arrival Delay Regression Model (Open Ended Variation)"
author: "[Kevin Jue]"
date: "[4/18/2017]"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYCFlights Model

Using the rectangular data that you created last week create a model for `arr_delay`. Follow a well defined process for creating a model as discussed in class or from one of the readings. 

Break the process down into steps and create a codeblock for each step. 
 - Use as many steps you feel as necessary
 - Concisely explain each step 
 - Show all work
 
(You can start with the data that you created last week.)


## Step 1: [Create Data] 

```{r "Step 1: Setup/Read/Transform Dataset" }
#install the packages in order to complete the operation
library(readr)
library(magrittr)
library(dplyr)
library(data.table)
library(ggplot2)

#read the data into R
setwd('~/csx460/02-building-blocks/02-exercise-nycflights/data/')
flights <- setDT(read.csv('flights.csv'))
planes <- setDT(read.csv('planes.csv'))
airports <- setDT(read.csv('airports.csv'))
weather <- setDT(read.csv('weather.csv'))

#perform the necessary joins in order to make a rectangular data table

flights_join <- left_join(flights, planes, by = 'tailnum', suffix = c(".of_flights",".of_planes")) %>%
                left_join(weather, by = c('origin', 'time_hour'), suffix = c(".of_flights",".of_weather")) %>%
                left_join(airports, by = c('origin' = 'faa'), suffix = c(".of_flights",".of_origin")) %>%
                left_join(airports, by = c('dest' = 'faa'), suffix = c(".of_origin",".of_dest"))


```

## Step 2: Create Naive Model
```{r "Step 2: Set a baseline of model performance by calculating the mean and the SSE of it"}

arr_delay_mean <- mean(flights_join$arr_delay, na.rm = TRUE)
SSE_mean <- sum((flights_join$arr_delay - arr_delay_mean)^2, na.rm = TRUE)

```


## Step 3: Calculate a base linear regression
```{r "Step 3: As a first step run the simplest linear regression with all variables as predictors"}

base_lm <- lm(arr_delay ~ month.of_flights + day.of_flights +	dep_time + sched_dep_time + sched_arr_time + flight + tailnum	+ origin + dest +	air_time	+ distance +	hour.of_flights + minute +	time_hour +	year.of_planes + engines + seats + speed + month.of_weather + day.of_weather + hour.of_weather + temp + dewp + humid + wind_dir +	wind_speed + wind_gust + precip + pressure + visib +	name.of_origin + lat.of_origin + lon.of_origin + alt.of_origin + tz.of_origin + name.of_dest + lat.of_dest + lon.of_dest + alt.of_dest + tz.of_dest + dst.of_dest +year.of_flights + year, data = flights_join,  na.rm = TRUE)


SSE <- sum((flights_join$arr_delay - arr_delay_mean)^2, na.rm = TRUE)

```


## Step 4: Compare the two models
```{r "Step 4: Runs SSE for both models and see which has a lower value, this should be a very basic test for determining which model is a better fit}

#if i could get the linear model in step 3 to run without having a memory overload, we would look at the values with the highest correlation, and use those to create a new linear regression, leaving the less/non correlated values out.  Then we would check the new model using the SSE test we created to see how it compares against the naive model (the mean).

```


# Question:

Is this a good model?  (Write your answer here.)

Step 4 would determine this.  If it can handily beat the naive model I would say it is an improvement.